---
title: "GitHub"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
date: "`r format(Sys.time(), '%d %B, %Y')`"

---

## Preview

```{r, echo = FALSE, hide = TRUE, message=FALSE, warning = FALSE}
library(ggplot2)
library(magrittr)
library(dplyr)

root_dir <- rprojroot::find_root(rprojroot::has_dir(".git"))
yaml <- yaml::read_yaml(file.path(root_dir, "_config_automation.yml"))
target_repos <- yaml$github_repos

## For github
github_overall <- readr::read_tsv(file.path("metricminer_data", "github", "github.tsv"))
github_timecourse <- readr::read_tsv(file.path("metricminer_data", "github", "github_timecourse.tsv"))

## For google
# github_overall <- googlesheets4::read_sheet(yaml$github_googlesheet, sheet = "overall_stats")
# github_timecourse <- googlesheets4::read_sheet(yaml$github_googlesheet, sheet = "timecourse")

github_overall <- github_overall %>%
  dplyr::filter(repo_name %in% target_repos)

github_timecourse <- github_timecourse %>%
  dplyr::filter(repo %in% target_repos) %>%
  dplyr::mutate(timestamp = as.Date(timestamp))

repo_label <- paste(target_repos, collapse = ", ")
has_github_data <- nrow(github_overall) > 0 || nrow(github_timecourse) > 0
```

This page summarizes GitHub repository activity for `r repo_label`.

```{r, echo = FALSE, results = "asis"}
if (!has_github_data) {
  cat(
    "No refreshed GitHub metric rows were found for `",
    repo_label,
    "` in `metricminer_data/github/` yet. ",
    "Run `Rscript refresh-scripts/refresh-github.R` and re-render this page after the refresh completes.",
    sep = ""
  )
}
```

Overall repository metrics:

```{r echo = FALSE, hide = TRUE, message=FALSE, warning = FALSE}
if (nrow(github_overall) > 0) {
  github_overall %>%
    tidyr::pivot_longer(
      cols = c(num_forks, num_contributors, total_contributions, num_stars),
      names_to = "metric",
      values_to = "value"
    ) %>%
    dplyr::filter(!is.na(value)) %>%
    ggplot(aes(x = metric, y = value)) +
    geom_col(fill = "#4C78A8") +
    facet_wrap(~ repo_name) +
    scale_x_discrete(
      labels = c(
        num_forks = "Forks",
        num_contributors = "Contributors",
        total_contributions = "Contributions",
        num_stars = "Stars"
      )
    ) +
    labs(x = NULL, y = "Count") +
    theme_classic()
}
```

Views and clones over time:

```{r echo = FALSE, hide = TRUE, message=FALSE}
if (nrow(github_timecourse) > 0) {
  github_timecourse %>%
    dplyr::select(repo, timestamp, count_clones, count_views) %>%
    tidyr::pivot_longer(
      cols = c(count_clones, count_views),
      names_to = "metric",
      values_to = "count"
    ) %>%
    dplyr::filter(!is.na(count)) %>%
    ggplot(aes(x = timestamp, y = count, color = metric)) +
    geom_line(linewidth = 1) +
    geom_point(size = 1.8) +
    facet_wrap(~ repo, scales = "free_y") +
    scale_color_manual(
      values = c(count_clones = "#F58518", count_views = "#4C78A8"),
      labels = c(count_clones = "Clones", count_views = "Views")
    ) +
    labs(x = "Date", y = "Count", color = NULL) +
    theme_classic()
}
```

## Data information 

Data information for GitHub overall data:

- `num_forks` shows the number of times this repo has been forked. NA means it has never been forked. 
- `num_contributors` how many people have contributed to this repo
- `num_stars` how many people have starred this repo?
- `health_percentage` what percentage of ["good software health" items as described by GitHub](https://docs.github.com/en/communities/setting-up-your-project-for-healthy-contributions/about-community-profiles-for-public-repositories) does this repo have? 

```{r, message=FALSE}
if (nrow(github_overall) > 0) {
  knitr::kable(github_overall)
}
```

Data information for GitHub timecourse data:

- `timestamp` shows the date the counts correspond to
- `count_clones` tells the number of clones made on this day. NAs generally indicate no one cloned the repository that day. 
- `uniques_clones` tells how many people did these clones come from? NAs generally indicate no one cloned the repository that day.
- `count_views` how many views did the repo get on this day? NAs generally indicate no one viewed the repository that date
- `uniques_views` how many people were those views from? NAs generally indicate no one viewed the repository that date

```{r, message=FALSE}
if (nrow(github_timecourse) > 0) {
  knitr::kable(github_timecourse)
}
```

## Setting up GitHub

This dashboard is configured to collect GitHub metrics for `BioDataSage/fair25`.

In the `_config_automation.yml` you will need to specify items in the GitHub section.
```
###### GitHub ######
refresh-github: yes
github_repos: [
  BioDataSage/fair25
  ]
github_googlesheet:
```

- [ ] In the `_config_automation.yml` file, make sure that `refresh-github` is set to "yes".
- [ ] In `github_repos` of your `_config_automation.yml`, keep `BioDataSage/fair25` listed using the `owner/repository` format.
- [ ] Optionally, if you are saving data to Google, specify a googlesheet ID in `github_googlesheet` you'd like the GitHub data to be saved to. This will only be relevant if you've set `data_dest` to `google`.

## Customizing GitHub Data 

In order to customize the data you are downloading from GitHub you can modify the 
`refresh-scripts/refresh-github.R` script in your repository. 

You can take a look at the [`metricminer` R package documentation](https://hutchdatascience.org/metricminer/articles/getting-started.html) for more details about the functions and what is possible. 

If you have a metric need that is not currently fulfilled by `metricminer` or `metricminer-dashboard` we encourage you to [file a GitHub issue with us and let us know about your new feature idea (or bug report)](https://github.com/fhdsl/metricminer/issues/new/choose). 
