---
title: "Google Forms"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

## Dashboard

```{r, echo = FALSE, hide = TRUE, message = FALSE}
root_dir <- rprojroot::find_root(rprojroot::has_dir(".git"))
yaml <- yaml::read_yaml(file.path(root_dir, "_config_automation.yml"))

library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(scales)
library(readr)
library(purrr)

`%||%` <- function(x, y) {
  if (is.null(x)) y else x
}

googleforms_dir <- file.path("metricminer_data", "googleforms")

get_googleforms_sheet_id <- function(yaml) {
  yaml$googleforms_googlesheet %||% yaml$gf_googlesheet
}

has_value <- function(x) {
  !is.null(x) && length(x) > 0 && !is.na(x) && x != ""
}

read_googleforms_metric <- function(path, sheet) {
  if (file.exists(path)) {
    return(readr::read_tsv(path, show_col_types = FALSE))
  }

  sheet_id <- get_googleforms_sheet_id(yaml)
  if (has_value(sheet_id)) {
    google_data <- try(
      googlesheets4::read_sheet(sheet_id, sheet = sheet),
      silent = TRUE
    )

    if (!inherits(google_data, "try-error")) {
      return(as_tibble(google_data))
    }
  }

  NULL
}

score_answer <- function(answer) {
  parsed <- suppressWarnings(readr::parse_number(as.character(answer)))
  likert_scores <- c(
    "strongly disagree" = 1,
    "disagree" = 2,
    "neutral" = 3,
    "neither agree nor disagree" = 3,
    "agree" = 4,
    "strongly agree" = 5,
    "not at all confident" = 1,
    "slightly confident" = 2,
    "moderately confident" = 3,
    "very confident" = 4,
    "extremely confident" = 5,
    "novice" = 1,
    "beginner" = 2,
    "intermediate" = 3,
    "advanced" = 4,
    "expert" = 5
  )

  dplyr::coalesce(parsed, unname(likert_scores[str_squish(str_to_lower(answer))]))
}

build_fallback_responses <- function() {
  if (!identical(yaml$data_dest, "github")) {
    return(tibble())
  }

  sample_files <- list.files(googleforms_dir, pattern = "\\.tsv$", full.names = TRUE)
  sample_files <- sample_files[!basename(sample_files) %in% c(
    "survey_responses_long.tsv",
    "survey_question_metrics.tsv",
    "survey_answer_metrics.tsv",
    "survey_change_metrics.tsv",
    "survey_growth_metrics.tsv"
  )]

  if (length(sample_files) == 0) {
    return(tibble())
  }

  stages <- yaml$google_form_stages
  if (is.null(stages) || length(stages) < length(sample_files)) {
    stages <- c(stages, c("pre", "post")[seq_len(length(sample_files))])
  }

  purrr::map2_dfr(sample_files, seq_along(sample_files), function(path, i) {
    readr::read_tsv(path, show_col_types = FALSE) %>%
      distinct() %>%
      pivot_longer(
        cols = ends_with("_answers"),
        names_to = "question_id",
        values_to = "answer",
        values_drop_na = TRUE
      ) %>%
      mutate(
        survey_stage = janitor::make_clean_names(stages[[i]]),
        survey_name = tools::file_path_sans_ext(basename(path)),
        question_id = str_remove(question_id, "_answers$"),
        question = question_id,
        numeric_score = score_answer(answer),
        participant_id = response_id
      ) %>%
      select(survey_stage, survey_name, response_id, participant_id, question_id, question, answer, numeric_score)
  })
}

configured_form_ids <- yaml$google_forms %||% character()
configured_form_stages <- janitor::make_clean_names(yaml$google_form_stages %||% character())
configured_form_count <- max(length(configured_form_ids), length(configured_form_stages))

if (configured_form_count == 0) {
  configured_forms <- tibble(
    form_stage = character(),
    google_form_id = character()
  )
} else {
  configured_forms <- tibble(
    form_stage = c(configured_form_stages, rep(NA_character_, configured_form_count))[seq_len(configured_form_count)],
    google_form_id = c(configured_form_ids, rep(NA_character_, configured_form_count))[seq_len(configured_form_count)]
  )
}

googleforms_sheet_id <- get_googleforms_sheet_id(yaml)

empty_responses <- function() {
  tibble(
    survey_stage = character(),
    survey_name = character(),
    response_id = character(),
    participant_id = character(),
    question_id = character(),
    question = character(),
    answer = character(),
    numeric_score = double()
  )
}

empty_question_metrics <- function() {
  tibble(
    survey_stage = character(),
    survey_name = character(),
    question_id = character(),
    question = character(),
    responses = integer(),
    numeric_responses = integer(),
    mean_score = double(),
    median_score = double()
  )
}

empty_answer_metrics <- function() {
  tibble(
    survey_stage = character(),
    survey_name = character(),
    question_id = character(),
    question = character(),
    answer = character(),
    responses = integer(),
    total_responses = integer(),
    proportion = double()
  )
}

empty_growth_metrics <- function() {
  tibble(
    participant_id = character(),
    question_id = character(),
    question = character(),
    question_key = character(),
    pre_answer = character(),
    post_answer = character(),
    answer_changed = logical(),
    pre_score = double(),
    post_score = double(),
    score_change = double(),
    percent_score_change = double()
  )
}

empty_change_metrics <- function() {
  tibble(
    question_id = character(),
    question = character(),
    pre_answer = character(),
    post_answer = character(),
    participants = integer(),
    proportion = double()
  )
}

responses_path <- file.path(googleforms_dir, "survey_responses_long.tsv")
question_metrics_path <- file.path(googleforms_dir, "survey_question_metrics.tsv")
answer_metrics_path <- file.path(googleforms_dir, "survey_answer_metrics.tsv")
growth_metrics_path <- file.path(googleforms_dir, "survey_growth_metrics.tsv")
change_metrics_path <- file.path(googleforms_dir, "survey_change_metrics.tsv")

survey_responses <- read_googleforms_metric(responses_path, "survey_responses_long")

survey_responses <- if (is.null(survey_responses)) {
  build_fallback_responses()
} else {
  survey_responses
}

if (nrow(survey_responses) == 0) {
  survey_responses <- empty_responses()
}

survey_responses <- survey_responses %>%
  mutate(
    survey_stage = factor(survey_stage, levels = c("pre", "post")),
    question = stringr::str_wrap(question, width = 48)
  )

survey_question_metrics <- read_googleforms_metric(question_metrics_path, "survey_question_metrics")

survey_question_metrics <- if (is.null(survey_question_metrics) && nrow(survey_responses) > 0) {
  survey_responses %>%
    filter(!is.na(answer)) %>%
    group_by(survey_stage, survey_name, question_id, question) %>%
    summarise(
      responses = n_distinct(response_id),
      numeric_responses = sum(!is.na(numeric_score)),
      mean_score = if_else(numeric_responses > 0, mean(numeric_score, na.rm = TRUE), NA_real_),
      median_score = if_else(numeric_responses > 0, median(numeric_score, na.rm = TRUE), NA_real_),
      .groups = "drop"
    )
} else if (is.null(survey_question_metrics)) {
  empty_question_metrics()
} else {
  survey_question_metrics
}

survey_question_metrics <- survey_question_metrics %>%
  mutate(
    survey_stage = factor(survey_stage, levels = c("pre", "post")),
    question = stringr::str_wrap(question, width = 48)
  )

survey_answer_metrics <- read_googleforms_metric(answer_metrics_path, "survey_answer_metrics")

survey_answer_metrics <- if (is.null(survey_answer_metrics) && nrow(survey_responses) > 0) {
  survey_responses %>%
    filter(!is.na(answer)) %>%
    group_by(survey_stage, survey_name, question_id, question, answer) %>%
    summarise(responses = n_distinct(response_id), .groups = "drop_last") %>%
    mutate(
      total_responses = sum(responses),
      proportion = responses / total_responses
    ) %>%
    ungroup()
} else if (is.null(survey_answer_metrics)) {
  empty_answer_metrics()
} else {
  survey_answer_metrics
}

survey_answer_metrics <- survey_answer_metrics %>%
  mutate(
    survey_stage = factor(survey_stage, levels = c("pre", "post")),
    question = stringr::str_wrap(question, width = 48),
    answer = stringr::str_wrap(answer, width = 32)
  )

survey_growth <- read_googleforms_metric(growth_metrics_path, "survey_growth_metrics")

survey_growth <- if (is.null(survey_growth)) {
  empty_growth_metrics()
} else {
  survey_growth
}

survey_change_metrics <- read_googleforms_metric(change_metrics_path, "survey_change_metrics")

survey_change_metrics <- if (is.null(survey_change_metrics) && nrow(survey_growth) > 0) {
  survey_growth %>%
    group_by(question_id, question, pre_answer, post_answer) %>%
    summarise(participants = n_distinct(participant_id), .groups = "drop_last") %>%
    mutate(proportion = participants / sum(participants)) %>%
    ungroup()
} else if (is.null(survey_change_metrics)) {
  empty_change_metrics()
} else {
  survey_change_metrics
}

survey_change_metrics <- survey_change_metrics %>%
  mutate(
    question = stringr::str_wrap(question, width = 48),
    answer_change = stringr::str_wrap(paste(pre_answer, "to", post_answer), width = 44)
  )

stage_summary <- survey_responses %>%
  filter(!is.na(survey_stage)) %>%
  group_by(survey_stage) %>%
  summarise(
    responses = n_distinct(response_id),
    participants = n_distinct(participant_id),
    answers = sum(!is.na(answer)),
    .groups = "drop"
  )

```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::kable(configured_forms, caption = "Configured Google Forms")
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
data_source <- if (file.exists(responses_path)) {
  "local TSV data"
} else if (has_value(googleforms_sheet_id)) {
  paste0("Google Sheet: ", googleforms_sheet_id)
} else {
  "not populated yet"
}

dashboard_status <- tibble(
  setting = c("data destination", "dashboard data source"),
  value = c(yaml$data_dest %||% "not configured", data_source)
)

knitr::kable(dashboard_status, caption = "Dashboard status")
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
if (nrow(stage_summary) > 0) {
  knitr::kable(stage_summary, caption = "Survey completion summary")
} else {
  knitr::asis_output("No Google Forms response data is available yet. Run `refresh-scripts/refresh-googleforms.R` after Google authentication is configured to create or populate the dashboard data sheet.")
}
```

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 5}
if (nrow(survey_answer_metrics) > 0) {
  survey_answer_metrics %>%
    group_by(question, survey_stage) %>%
    slice_max(order_by = responses, n = 8, with_ties = FALSE) %>%
    ungroup() %>%
    ggplot(aes(x = proportion, y = answer, fill = survey_stage)) +
    geom_col(position = position_dodge2(preserve = "single"), width = 0.7) +
    facet_wrap(~question, scales = "free_y") +
    scale_fill_manual(values = c(pre = "#2f6f73", post = "#d85c3a")) +
    scale_x_continuous(labels = scales::percent_format()) +
    labs(
      title = "Answer distribution before and after the workshop",
      x = "Share of responses",
      y = NULL,
      fill = NULL
    ) +
    theme_minimal(base_size = 12) +
    theme(panel.grid.minor = element_blank())
} else {
  knitr::asis_output("No survey answers were found yet.")
}
```

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4}
if (nrow(survey_change_metrics) > 0) {
  survey_change_metrics %>%
    group_by(question) %>%
    slice_max(order_by = participants, n = 10, with_ties = FALSE) %>%
    ungroup() %>%
    ggplot(aes(x = participants, y = reorder(answer_change, participants))) +
    geom_col(fill = "#2f6f73", width = 0.7) +
    facet_wrap(~question, scales = "free_y") +
    labs(
      title = "How matched participants changed their answers",
      x = "Participants",
      y = NULL
    ) +
    theme_minimal(base_size = 12) +
    theme(panel.grid.minor = element_blank())
} else {
  knitr::asis_output("Paired growth metrics will appear after pre and post responses can be matched by participant.")
}
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
if (nrow(survey_growth) > 0) {
  survey_growth %>%
    group_by(question) %>%
    summarise(
      paired_participants = n_distinct(participant_id),
      changed_answers = sum(answer_changed, na.rm = TRUE),
      percent_changed = changed_answers / paired_participants,
      .groups = "drop"
    ) %>%
    arrange(desc(percent_changed)) %>%
    knitr::kable(digits = 2, caption = "Paired pre/post answer changes")
}
```

## Google Forms Configuration

[Follow the steps from the above section to authenticate Google](#setting-up-google-authentication) -- make sure that the account you have authenticated for has access to the Google Forms you wish to collect.

Go to a form you'd like to collect data from. The URL should look something like this:
```
https://docs.google.com/forms/u/1/d/<GOOGLE_FORM_ID_HERE>/edit?usp=drive_web
```
Extract the Google Form ID from the URL and put it in the `google_forms` section of the `_config_automation.yml` file.
Do this for each form you'd like to collect data from and separate the form IDs with commas. Delete the example IDs we have put there as placeholders.

```
###### Google Forms ######
refresh-googleforms: yes
google_forms: [
  1pbFfgUPYH2w9zEoCDjCa4HFOxzEhGOseufw28Xxmd-o,
  1JjmsiuVoGSxvl-1M_oWittcftO955tijzeNc-cgJlo8 ]
google_form_stages: [ pre, post ]
googleforms_participant_id_question:
googleforms_googlesheet:
```

- [ ] In the `_config_automation.yml` file, make sure that `refresh-googleforms` is set to "yes".
- [ ] Put the pre-workshop survey first and the post-workshop survey second in `google_forms`. Keep `google_form_stages` in the same order.
- [ ] If the dashboard cannot infer the participant matching question, set `googleforms_participant_id_question` to the exact question title or question ID used for participant email, name, or ID.
- [ ] Optionally, if you are saving data to Google, specify a googlesheet ID in `googleforms_googlesheet` you'd like the Google Forms data to be saved to. If this is blank, the refresh script will create a sheet after Google authentication is configured.

## Customizing the Google Forms Data

In order to customize the data you are downloading from your Google Forms you can modify the
`refresh-scripts/refresh-googleforms.R` script in your repository.

You can take a look at the [`metricminer` R package documentation](https://hutchdatascience.org/metricminer/articles/getting-started.html) for more details about the functions and what is possible.

If you have a metric need that is not currently fulfilled by `metricminer` or `metricminer-dashboard` we encourage you to [file a GitHub issue with us and let us know about your new feature idea (or bug report)](https://github.com/fhdsl/metricminer/issues/new/choose).
